---
output:
  rmarkdown::github_document
bibliography: "inst/REFERENCES.bib"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`tmle3shift`

[![Travis-CI Build Status](https://travis-ci.org/tlverse/tmle3shift.svg?branch=master)](https://travis-ci.org/tlverse/tmle3shift)
[![AppVeyor Build  Status](https://ci.appveyor.com/api/projects/status/github/tlverse/tmle3shift?branch=master&svg=true)](https://ci.appveyor.com/project/nhejazi/tmle3shift)
[![Coverage Status](https://img.shields.io/codecov/c/github/tlverse/tmle3shift/master.svg)](https://codecov.io/github/tlverse/tmle3shift?branch=master)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![License: GPL v3](https://img.shields.io/badge/License-GPL%20v3-blue.svg)](http://www.gnu.org/licenses/gpl-3.0)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.4603372.svg)](https://doi.org/10.5281/zenodo.4603372)

> Targeted Learning of the Causal Effects of Stochastic Interventions

__Authors:__ [Nima Hejazi](https://nimahejazi.org), [Jeremy
Coyle](https://github.com/jeremyrcoyle), and [Mark van der
Laan](https://vanderlaan-lab.org)

---

## What's `tmle3shift`?

`tmle3shift` is an adapter/extension R package in the `tlverse` ecosystem that
exposes support for the estimation of a target parameter defined as the mean
counterfactual outcome under a posited shift of the natural value of a
continuous-valued intervention, using the formalism of stochastic treatment
regimes. As an adapter package, `tmle3shift` builds upon the core `tlverse`
grammar introduced by `tmle3`, a general framework that supports the
implementation of a range of TMLE parameters through a unified interface. For a
detailed description of the target parameter, TML estimator, and algorithm
implemented in `tmle3shift`, the interested reader is invited to consult
@diaz2012population and @diaz2018stochastic. For a general discussion of the
framework of targeted minimum loss-based estimation and the role this
methodology plays in statistical and causal inference, the canonical references
are @vdl2011targeted and @vdl2018targeted.

Building on the original work surrounding the TML estimator for the
aforementioned target parameter, `tmle3shift` additionally implements a set of
techniques for variable importance analysis, allowing for a sequence of mean
counterfactual outcomes, estimated under a sequence of posited shifts, to be
summarized via a working marginal structural model (MSM). The goal of this work
is to build upon the `tlverse` framework and the estimation methodology
implemented for a single mean counterfactual outcome in order to introduce an
end-to-end methodology for variable importance analyses.

---

## Installation

You can install the development version of `tmle3shift` from GitHub via
[`remotes`](https://CRAN.R-project.org/package=remotes) with

```{r gh-installation, eval = FALSE}
remotes::install_github("tlverse/tmle3shift")
```

---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/tlverse/tmle3shift/issues).

---

## Contributions

Contributions are very welcome. Interested contributors should consult our
[contribution
guidelines](https://github.com/tlverse/tmle3shift/blob/master/CONTRIBUTING.md)
prior to submitting a pull request.

---

## Citation

After using the `tmle3shift` R package, please cite the following:

        @software{hejazi2021tmle3shift-rpkg,
          author = {Hejazi, Nima S and Coyle, Jeremy R and {van der Laan}, Mark
            J},
          title = {{tmle3shift}: {Targeted Learning} of the Causal Effects of
            Stochastic Interventions},
          year = {2021},
          howpublished = {\url{https://github.com/tlverse/tmle3shift}},
          note = {{R} package version 0.2.0},
          url = {https://doi.org/10.5281/zenodo.4603372},
          doi = {10.5281/zenodo.4603372}
        }

---

## Related

* [R/`txshift`](https://github.com/nhejazi/txshift) - An R package providing an
  independent implementation of the TML estimation procedure and
  statistical methodology as is made available here, without reliance on the
  `tlverse` grammar provided by `tmle3`.

---

## Funding

The development of this software was supported in part through a grant from the
National Institutes of Health: [T32
LM012417-02](https://projectreporter.nih.gov/project_info_description.cfm?aid=9248418&icde=37849831&ddparam=&ddvalue=&ddsub=&cr=1&csb=default&cs=ASC&pball=).

---

## License

The contents of this repository are distributed under the GPL-3 license. See
file `LICENSE` for details.

---

## References

