---
output: rmarkdown::github_document
bibliography: "`r here::here('README-refs.bib')`"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`tmle3shiftshift`

[![Travis-CI Build Status](https://travis-ci.org/tlverse/tmle3shift.svg?branch=master)](https://travis-ci.org/tlverse/tmle3shift)
[![AppVeyor Build  Status](https://ci.appveyor.com/api/projects/status/github/tlverse/tmle3shift?branch=master&svg=true)](https://ci.appveyor.com/project/tlverse/tmle3shift)
[![Coverage Status](https://img.shields.io/codecov/c/github/tlverse/tmle3shift/master.svg)](https://codecov.io/github/tlverse/tmle3shift?branch=master)
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)
[![License: GPL v3](https://img.shields.io/badge/License-GPL%20v3-blue.svg)](http://www.gnu.org/licenses/gpl-3.0)

> TML Estimation of Shift Intervention Effects with the `tmle3` framework

---

## What's `tmle3shift`?

`tmle3shift` is an adapter / plug-in R package for use with `tmle3`, a general
framework that supports the implementation of a range of TMLE parameters with a
unified interface. `tmle3shift` extends support to the estimation of parameters
defined as shift interventions of continuous-valued treatments by way of
stochastic treatment regimes. For a detailed description of the parameter, TML
estimator, and algorithm implemented in `tmle3shift`, the interested reader is
invited to consult @munoz2012population and @diaz2018stochastic.

For a general discussion of the framework of targeted minimum loss-based
estimation and the role this methodology plays in statistical causal inference,
the canonical references are @vdl2011targeted and @vdl2018targeted.

---

## Installation

You can install the development version of `tmle3shift` from GitHub via
[`devtools`](https://www.rstudio.com/products/rpackages/devtools/) with

```{r gh-installation, eval = FALSE}
devtools::install_github("tlverse/tmle3shift")
```

---

## Getting Started

The best place to get started is the "Framework Overview" document, which
describes the individual components of the `tmle3shift` framework. It may be found
here: https://tmle3shift.tlverse.org/articles/framework.html.

---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/tlverse/tmle3shift/issues).

---

## License

The contents of this repository are distributed under the GPL-3 license. See
file `LICENSE` for details.

---

## References

