---
output:
  rmarkdown::github_document
bibliography: "README-refs.bib"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`tmle3shift`

[![Travis-CI Build Status](https://travis-ci.org/tlverse/tmle3shift.svg?branch=master)](https://travis-ci.org/tlverse/tmle3shift)
[![AppVeyor Build  Status](https://ci.appveyor.com/api/projects/status/github/tlverse/tmle3shift?branch=master&svg=true)](https://ci.appveyor.com/project/tlverse/tmle3shift)
[![Coverage Status](https://img.shields.io/codecov/c/github/tlverse/tmle3shift/master.svg)](https://codecov.io/github/tlverse/tmle3shift?branch=master)
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)
[![License: GPL v3](https://img.shields.io/badge/License-GPL%20v3-blue.svg)](http://www.gnu.org/licenses/gpl-3.0)

> Targeted Learning of Stochastic Intervention Effects, done `tlverse`-style

---

## What's `tmle3shift`?

`tmle3shift` is an adapter/extension R package in the `tlverse` ecosystem that
exposes support for the estimation of target parameters defined as shifts of
interventions of continuous-valued treatments by way of stochastic treatment
regimes. As an adapter package, `tmle3shift` builds upon the core `tlverse`
grammar introduced by `tmle3`, a general framework that supports the
implementation of a range of TMLE parameters through a unified interface. For a
detailed description of the target parameter, TML estimator, and algorithm
implemented in `tmle3shift`, the interested reader is invited to consult
@munoz2012population and @diaz2018stochastic.

For a general discussion of the framework of targeted minimum loss-based
estimation and the role this methodology plays in statistical and causal
inference, the canonical references are @vdl2011targeted and @vdl2018targeted.

---

## Installation

You can install the development version of `tmle3shift` from GitHub via
[`devtools`](https://www.rstudio.com/products/rpackages/devtools/) with

```{r gh-installation, eval = FALSE}
devtools::install_github("tlverse/tmle3shift")
```

---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/tlverse/tmle3shift/issues).

---

## Related

* [R/`txshift`](https://github.com/nhejazi/txshift) - An R package providing an
  independent implementation of the same TML estimation procedure and
  statistical methodology as is made available here, without reliance on the
  `tlverse` grammar provided by the `tmle3` engine.

---

## Funding

The development of this software was supported in part through a grant from the
National Institutes of Health: [T32
LM012417-02](https://projectreporter.nih.gov/project_info_description.cfm?aid=9248418&icde=37849831&ddparam=&ddvalue=&ddsub=&cr=1&csb=default&cs=ASC&pball=).

---

## License

The contents of this repository are distributed under the GPL-3 license. See
file `LICENSE` for details.

---

## References

