---
title: "Targeted Learning with Stochastic Treatment Regimes"
author: "[Nima Hejazi](https://nimahejazi.org) and [Jeremy
  Coyle](https://github.com/jeremyrcoyle)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: vignette-refs.bib
vignette: >
  %\VignetteIndexEntry{Targeted Learning with Stochastic Treatment Regimes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r options, echo=FALSE}
options(scipen=999)
```

## Introduction

Stochastic treatment regimes present a relatively simple manner in which to
assess the effects of continuous treatments by way of parameters that examine
the effects induced by the counterfactual shifting of the observed values of a
treatment of interest. Here, we present an implementation of a new algorithm for
computing targeted minimum loss-based estimates of treatment shift parameters
defined based on a shifting function $d(A,W)$. For a technical presentation of
the algorithm, the interested reader is invited to consult @diaz2018stochastic.
For additional background on Targeted Learning and previous work on stochastic
treatment regimes, please consider consulting @vdl2011targeted,
@vdl2018targeted, and @diaz2012population.

To start, let's load the packages we'll use and set a seed for simulation:

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(sl3)
library(tmle3)
library(tmle3shift)
set.seed(429153)
```

## Data and Notation

Consider $n$ observed units $O_1, \ldots, O_n$, where each random variable $O =
(W, A, Y)$ corresponds to a single observational unit. Let $W$ denote baseline
covariates (e.g., age, sex, education level), $A$ an intervention variable of
interest (e.g., nutritional supplements), and $Y$ an outcome of interest (e.g.,
disease status). Though it need not be the case, let $A$ be continuous-valued,
i.e. $A \in \mathbb{R}$. Let $O_i \sim \mathcal{P} \in \mathcal{M}$, where
$\mathcal{M}$ is the nonparametric statistical model defined as the set of
continuous densities on $O$ with respect to some dominating measure. To
formalize the definition of stochastic interventions and their corresponding
causal effects, we introduce a nonparametric structural equation model (NPSEM),
based on @pearl2000causality, to define how the system changes under posited
interventions:
\begin{align*}\label{eqn:npsem}
  W &= f_W(U_W) \\ A &= f_A(W, U_A) \\ Y &= f_Y(A, W, U_Y),
\end{align*}
We denote the observed data structure $O = (W, A, Y)$

Letting $A$ denote a continuous-valued treatment, we assume that the
distribution of $A$ conditional on $W = w$ has support in the interval
$(l(w), u(w))$ -- for convenience, let this support be _a.e._ That is, the
minimum natural value of treatment $A$ for an individual with covariates
$W = w$ is $l(w)$; similarly, the maximum is $u(w)$. Then, a simple stochastic
intervention, based on a shift $\delta$, may be defined
\begin{equation}\label{eqn:shift}
  d(a, w) =
  \begin{cases}
    a - \delta & \text{if } a > l(w) + \delta \\
    a & \text{if } a \leq l(w) + \delta,
  \end{cases}
\end{equation}
where $0 \leq \delta \leq u(w)$ is an arbitrary pre-specified value that
defines the degree to which the observed value $A$ is to be shifted, where
possible. For the purpose of using such a shift in practice, the present
software provides the functions `shift_additive` and `shift_additive_inv`,
which define a variation of this shift, assuming that the density of treatment
$A$, conditional on the covariates $W$, has support _a.e._

### Simulate Data

```{r sim_data}
# simulate simple data for tmle-shift sketch
n_obs <- 1000 # number of observations
n_w <- 1 # number of baseline covariates
tx_mult <- 2 # multiplier for the effect of W = 1 on the treatment

## baseline covariates -- simple, binary
W <- as.numeric(replicate(n_w, rbinom(n_obs, 1, 0.5)))

## create treatment based on baseline W
A <- as.numeric(rnorm(n_obs, mean = tx_mult * W, sd = 1))

## create outcome as a linear function of A, W + white noise
Y <- A + W + rnorm(n_obs, mean = 0, sd = 0.5)
```

The above composes our observed data structure $O = (W, A, Y)$. To formally
express this fact using the `tlverse` grammar introduced by the `tmle3` package,
we create a single data object and specify the functional relationships between
the nodes in the _directed acyclic graph_ (DAG) via _nonparametric structural
equation models_ (NPSEMs), reflected in the node list that we set up:

```{r data_nodes}
# organize data and nodes for tmle3
data <- data.table(W, A, Y)
node_list <- list(W = "W", A = "A", Y = "Y")
head(data)
```

We now have an observed data structure (`data`) and a specification of the role
that each variable in the data set plays as the nodes in a DAG.

## Methodology

To start, we will initialize a specification for the TMLE of our parameter of
interest (called a `tmle3_Spec` in the `tlverse` nomenclature) simply by calling
`tmle_shift`. We specify the argument `shift_val = 0.5` when initializing the
`tmle3_Spec` object to communicate that we're interested in a shift of $0.5$ on
the scale of the treatment $A$ -- that is, we specify $\delta = 0.5$ (note that
this is an arbitrarily chosen value for this example).

```{r spec_init}
# initialize a tmle specification
tmle_spec <- tmle_shift(shift_val = 0.5,
                        shift_fxn = shift_additive,
                        shift_fxn_inv = shift_additive_inv)
```

As seen above, the `tmle_shift` specification object (like all `tmle3_Spec`
objects) does _not_ store the data for our specific analysis of interest. Later,
we'll see that passing a data object directly to the `tmle3` wrapper function,
alongside the instantiated `tmle_spec`, will serve to construct a `tmle3_Task`
object internally (see the `tmle3` documentation for details).

### _Interlude:_ Constructing Optimal Stacked Regressions with `sl3`

To easily incorporate ensemble machine learning into the estimation procedure,
we rely on the facilities provided in the [`sl3` R
package](https://tlverse.org/sl3). For a complete guide on using the `sl3` R
package, consider consulting https://tlverse.org/sl3, or https://tlverse.org for
the [`tlverse` ecosystem](https://github.com/tlverse), of which `sl3` is a major
part.

Using the framework provided by the [`sl3` package](https://tlverse.org/sl3),
the nuisance parameters of the TML estimator may be fit with ensemble learning,
using the cross-validation framework of the Super Learner algorithm of
@vdl2007super. To estimate the treatment mechanism (often denoted "g" in the
targeted learning literature), we must make use of learning algorithms
specifically suited to conditional density estimation; a list of such learners 
may be extracted from `sl3` by using `sl3_list_learners()`:

```{r sl3_density_lrnrs_search}
sl3_list_learners("density")
```

To proceed, we'll select two of the above learners, `Lrnr_haldensify` for using
the highly adaptive lasso for conditional density estimation, based on an
algorithm given by @diaz2011super, and `Lrnr_rfcde`, an approach for using
random forests for conditional density estimation [@popisil2018rfcde]:

```{r sl3_regression_lrnrs}
# learners used for conditional density regression (i.e., propensity score)
lrn_haldensify <- Lrnr_haldensify$new(
  n_bins = 5, grid_type = "equal_mass",
  lambda_seq = exp(seq(-1, -13, length = 500))
)
lrn_rfcde <- Lrnr_rfcde$new(
  n_trees = 1000, node_size = 5,
  n_basis = 31, output_type = "observed"
)
sl_lrn_dens <- Lrnr_sl$new(
  learners = list(lrn_haldensify, lrn_rfcde),
  metalearner = Lrnr_solnp_density$new()
)
```

We also required an approach for estimating the outcome regression (often
denoted "Q" in the targeted learning literature). For this, we build a Super
Learner composed of an intercept model, a GLM, and the [xgboost
algorithm](https://xgboost.ai/) for gradient boosting:

```{r sl3_density_lrnrs}
# learners used for conditional expectation regression (e.g., outcome)
lrn_mean <- Lrnr_mean$new()
lrn_fglm <- Lrnr_glm_fast$new()
lrn_xgb <- Lrnr_xgboost$new(nrounds = 10)
sl_lrn <- Lrnr_sl$new(
  learners = list(lrn_mean, lrn_fglm, lrn_xgb),
  metalearner = Lrnr_nnls$new()
)
```

We can make the above explicit with respect to standard notation by bundling
the ensemble learners into a `list` object below. Note that in making the
`learner_list` object for our TML estimator, for the sake of example, we use
only the conditional density estimation approach of the [`haldensify`
package](https://github.com/nhejazi/haldensify) to estimate the treatment
mechanism; as this is done in an effort to save computation time, we recommend
using a stacked regression approach in standard data analytic practice.

```{r make_lrnr_list}
# specify outcome and treatment regressions and create learner list
Q_learner <- sl_lrn
#g_learner <- sl_lrn_dens     # in practice
g_learner <- lrn_haldensify
learner_list <- list(Y = Q_learner, A = g_learner)
```

The `learner_list` object above specifies the role that each of the ensemble
learners we've generated is to play in computing initial estimators to be used
in building a TMLE for the parameter of interest here. In particular, it makes
explicit the fact that our `Q_learner` is used in fitting the outcome regression
while our `g_learner` is used in fitting our treatment mechanism regression.

### Targeted Estimation of Stochastic Interventions Effects

```{r fit_tmle}
tmle_fit <- tmle3(tmle_spec, data, node_list, learner_list)
tmle_fit
```

The `print` method of the resultant `tmle_fit` object conveniently displays the
results from computing our TML estimator.

### Statistical Inference for Targeted Maximum Likelihood Estimates

Recall that the asymptotic distribution of TML estimators has been studied
thoroughly:
$$\psi_n - \psi_0 = (P_n - P_0) \cdot D(\bar{Q}_n^*, g_n) + R(\hat{P}^*, P_0),$$
which, provided the following two conditions:

1. If $D(\bar{Q}_n^*, g_n)$ converges to $D(P_0)$ in $L_2(P_0)$ norm, and
2. the size of the class of functions considered for estimation of $\bar{Q}_n^*$
   and $g_n$ is bounded (technically, $\exists \mathcal{F}$ st
   $D(\bar{Q}_n^*, g_n) \in \mathcal{F}$ *__whp__*, where $\mathcal{F}$ is a
   Donsker class),
readily admits the conclusion that
$\psi_n - \psi_0 = (P_n - P_0) \cdot D(P_0) + R(\hat{P}^*, P_0)$.

Under the additional condition that the remainder term $R(\hat{P}^*, P_0)$
decays as $o_P \left( \frac{1}{\sqrt{n}} \right),$ we have that
$$\psi_n - \psi_0 = (P_n - P_0) \cdot D(P_0) + o_P \left( \frac{1}{\sqrt{n}}
 \right),$$
which, by a central limit theorem, establishes a Gaussian limiting distribution
for the estimator:

$$\sqrt{n}(\psi_n - \psi) \to N(0, V(D(P_0))),$$
where $V(D(P_0))$ is the variance of the efficient influence curve (canonical
gradient) when $\psi$ admits an asymptotically linear representation.

The above implies that $\psi_n$ is a $\sqrt{n}$-consistent estimator of $\psi$,
that it is asymptotically normal (as given above), and that it is locally
efficient. This allows us to build Wald-type confidence intervals in a
straightforward manner:

$$\psi_n \pm z_{\alpha} \cdot \frac{\sigma_n}{\sqrt{n}},$$
where $\sigma_n^2$ is an estimator of $V(D(P_0))$. The estimator $\sigma_n^2$
may be obtained using the bootstrap or computed directly via the following

$$\sigma_n^2 = \frac{1}{n} \sum_{i = 1}^{n} D^2(\bar{Q}_n^*, g_n)(O_i)$$

Having now re-examined these facts, let's simply examine the results of
computing our TML estimator:

```{r tmle_inference}
tmle_fit
```

## References

