---
title: "Targeted Learning with Stochastic Treatment Regimes"
author: "[Nima Hejazi](https://nimahejazi.org), [David
  Benkeser](https://www.benkeserstatistics.com/), and [Jeremy
  Coyle](https://github.com/jeremyrcoyle)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: vignette-refs.bib
vignette: >
  %\VignetteIndexEntry{Targeted Learning with Stochastic Treatment Regimes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
options(scipen=999)
```

## Introduction

Stochastic treatment regimes present a relatively simple manner in which to
assess the effects of continuous treatments by way of parameters that examine
the effects induced by the counterfactual shifting of the observed values of a
treatment of interest. Here, we present an implementation of a new algorithm for
computing targeted minimum loss-based estimates of treatment shift parameters
defined based on a shifting function $d(A,W)$. For a technical presentation of
the algorithm, the interested reader is invited to consult @diaz2018stochastic.
For additional background on Targeted Learning and previous work on stochastic
treatment regimes, please consider consulting @vdl2011targeted,
@vdl2018targeted, and @munoz2012population.

To start, let's load the packages we'll use and set a seed for simulation:

```{r setup}
library(tidyverse)
library(data.table)
library(condensier)
library(sl3)
library(txshift)
set.seed(429153)
```

---

## Data and Notation

1. Start with a simple additive shift -- i.e., $d(a,w) = a + \delta$ if
   $a < u(w) - \delta$ or $d(a, w) = a$ if $a \geq u(w) - \delta$.

2. The additive shift will have support everywhere (i.e.,
   $a < u(w)$ is true everywhere).

3. The data structure that we now observe is $O = (W, A, Y)$.

### Simulate Data

```{r}
# grab from unit test
```

---

## Methodology

```{r}
# basic tmle3 setup
```

```

### Interlude: Constructing Optimal Stacked Regressions with `sl3`

To easily incorporate ensemble machine learning into the estimation procedure,
we rely on the facilities provided in the [`sl3` R
package](https://sl3.tlverse.org). For a complete guide on using the `sl3` R
package, consider consulting https://sl3.tlverse.org, or https://tlverse.org for
the `tlverse` ecosystem, of which `sl3` is a major part.

```{r}
# SL learners to be used for most fits (e.g., IPCW, outcome regression)
lrnr_lib <- make_learner_stack("Lrnr_mean", "Lrnr_glm_fast", "Lrnr_ranger")
sl_lrn <- Lrnr_sl$new(learners = lrnr_lib, metalearner = Lrnr_nnls$new())

# SL learners for conditional densities to be used for the propensity score fit
lrnr_dens_lib <- make_learner_stack(list("Lrnr_condensier", nbins = 35,
                                          bin_estimator = Lrnr_glm_fast$new(),
                                          bin_method = "equal.len",
                                          pool = FALSE),
                                    list("Lrnr_condensier", nbins = 30,
                                          bin_estimator = Lrnr_ranger$new(),
                                          bin_method = "equal.len",
                                          pool = FALSE),
                                    list("Lrnr_condensier", nbins = 25,
                                          bin_estimator = Lrnr_ranger$new(),
                                          bin_method = "equal.len",
                                          pool = FALSE)
                                   )
sl_lrn_dens <- Lrnr_sl$new(learners = lrnr_dens_lib,
                           metalearner = Lrnr_solnp_density$new())
```

### Estimating Stochastic Interventions Effects with Optimal Stacked Regressions

Using the framework provided by the [`sl3` package](https://sl3.tlverse.org),
the nuisance parameters of the TML estimator may be fit with ensemble learning,
using the cross-validation framework of the Super Learner algorithm of
@vdl2007super.

```{r}
# simple tmle3 fit
```

```{r}
# ???
```

### Statistical Inference for Targeted Maximum Likelihood Estimates

Recall that the asymptotic distribution of TML estimators has been studied
thoroughly:
$$\psi_n - \psi_0 = (P_n - P_0) \cdot D(\bar{Q}_n^*, g_n) + R(\hat{P}^*, P_0),$$
which, provided the following two conditions:

1. If $D(\bar{Q}_n^*, g_n)$ converges to $D(P_0)$ in $L_2(P_0)$ norm, and
2. the size of the class of functions considered for estimation of $\bar{Q}_n^*$
   and $g_n$ is bounded (technically, $\exists \mathcal{F}$ st
   $D(\bar{Q}_n^*, g_n) \in \mathcal{F}$ *__whp__*, where $\mathcal{F}$ is a
   Donsker class),
readily admits the conclusion that
$\psi_n - \psi_0 = (P_n - P_0) \cdot D(P_0) + R(\hat{P}^*, P_0)$.

Under the additional condition that the remainder term $R(\hat{P}^*, P_0)$
decays as $o_P \left( \frac{1}{\sqrt{n}} \right),$ we have that
$$\psi_n - \psi_0 = (P_n - P_0) \cdot D(P_0) + o_P \left( \frac{1}{\sqrt{n}}
 \right),$$
which, by a central limit theorem, establishes a Gaussian limiting distribution
for the estimator:

$$\sqrt{n}(\psi_n - \psi) \to N(0, V(D(P_0))),$$
where $V(D(P_0))$ is the variance of the efficient influence curve (canonical
gradient) when $\psi$ admits an asymptotically linear representation.

The above implies that $\psi_n$ is a $\sqrt{n}$-consistent estimator of $\psi$,
that it is asymptotically normal (as given above), and that it is locally
efficient. This allows us to build Wald-type confidence intervals in a
straightforward manner:

$$\psi_n \pm z_{\alpha} \cdot \frac{\sigma_n}{\sqrt{n}},$$
where $\sigma_n^2$ is an estimator of $V(D(P_0))$. The estimator $\sigma_n^2$
may be obtained using the bootstrap or computed directly via the following

$$\sigma_n^2 = \frac{1}{n} \sum_{i = 1}^{n} D^2(\bar{Q}_n^*, g_n)(O_i)$$

```{r}
# summary method for tmle3 fit object
```

## References

