---
title: "Variable Importance with Grids of Stochastic Interventions"
author: "[Nima S. Hejazi](https://nimahejazi.org), [Jeremy R.
  Coyle](https://github.com/jeremyrcoyle), and [Mark J. van der
  Laan](https://vanderlaan-lab.org)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: vignette-refs.bib
vignette: >
  %\VignetteIndexEntry{Variable Importance with Grids of Stochastic Interventions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r options, echo=FALSE, eval=FALSE}
options(scipen=999)
```

## Introduction

Stochastic treatment regimes present a relatively simple manner in which to
assess the effects of continuous treatments by way of parameters that examine
the effects induced by the counterfactual shifting of the observed values of a
treatment of interest. Here, we present an implementation of a new algorithm for
computing targeted minimum loss-based estimates of treatment shift parameters
defined based on a shifting function $d(A,W)$. For a technical presentation of
the algorithm, the interested reader is invited to consult @diaz2018stochastic.
For additional background on Targeted Learning and previous work on stochastic
treatment regimes, please consider consulting @vdl2011targeted,
@vdl2018targeted, and @diaz2012population.

To start, let's load the packages we'll use and set a seed for simulation:

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(condensier)
library(ranger)
library(sl3)
library(tmle3)
library(tmle3shift)
set.seed(429153)
```

---

## Data and Notation

1. Start with a simple additive shift -- i.e., $d(a,w) = a + \delta$ if
   $a leq u(w) - \delta$ or $d(a, w) = a$ if $a \geq u(w) - \delta$.

2. The additive shift will have support everywhere (i.e., $a \leq u(w)$ is true
   everywhere).

3. The data structure that we now observe is $O = (W, A, Y)$.

### Simulate Data

```{r sim_data}
# simulate simple data for tmle-shift sketch
n_obs <- 1000 # number of observations
n_w <- 1 # number of baseline covariates
tx_mult <- 2 # multiplier for the effect of W = 1 on the treatment

# baseline covariates -- simple, binary
W <- as.numeric(replicate(n_w, rbinom(n_obs, 1, 0.5)))

# create treatment based on baseline W
A <- as.numeric(rnorm(n_obs, mean = tx_mult * W, sd = 1))

# create outcome as a linear function of A, W + white noise
Y <- A + W + rnorm(n_obs, mean = 0, sd = 1)
```

The above composes our observed data structure $O = (W, A, Y)$. To formally
express this fact using the `tlverse` grammar introduced by the `tmle3` package,
we create a single data object and specify the functional relationships between
the nodes in the _directed acyclic graph_ (DAG) via _nonparametric structural
equation models_ (NPSEMs), reflected in the node list that we set up:

```{r data_nodes}
# organize data and nodes for tmle3
data <- data.table(W, A, Y)
node_list <- list(W = "W", A = "A", Y = "Y")
head(data)
```

We now have an observed data structure (`data`) and a specification of the role
that each variable in the data set plays as the nodes in a DAG.

---

## Methodology

### Defining a grid of counterfactual interventions

In order to specify a _grid_ of shifts $\delta$ to be used in defining a set of
stochastic intervention policies in an _a priori_ manner, let us consider an
arbitrary scalar $\delta$ that defines a counterfactual outcome $\psi_n =
Q_n(d(A, W), W)$, where, for simplicity, let $d(A, W) = A + \delta$. A
simplified expression of the auxiliary covariate for the TMLE of $\psi$ is
$H_n = \frac{g^*(a \mid w)}{g(a \mid w)}$, where $g^*(a \mid w)$ defines the
treatment mechanism with the stochastic intervention implemented. Then, to
ascertain whether a given choice of the shift $\delta$ is admissable (in the
sense of avoiding violations of the positivity assumption), let there be a bound
$C(\delta) = \frac{g^*(a \mid w)}{g(a \mid w)} < M$, where $g^*(a \mid w)$ is a
function of $\delta$ in part, and $M$ is a potentially user-specified upper
bound of $C(\delta)$. Then, $C(\delta)$ is a measure of the influence of a given
observation (under a bound of the conditional densities), which provides a way
to limit the maximum influence of a given observation through a choice of the
shift $\delta$.

We formalize and extend the procedure to determine an acceptable set of values
for the shift $\delta$ in the sequel. Specifically, let there be a shift $d(A,
W) = A + \delta(A, W)$, where the shift $\delta(A, W)$ is defined as
\begin{equation}
  \delta(a, w) =
    \begin{cases}
      \delta, & \delta_{\text{min}}(a,w) \leq \delta \leq
        \delta_{\text{max}}(a,w) \\
      \delta_{\text{max}}(a,w), & \delta \geq \delta_{\text{max}}(a,w) \\
      \delta_{\text{min}}(a,w), & \delta \leq \delta_{\text{min}}(a,w) \\
    \end{cases},
\end{equation}
where $$\delta_{\text{max}}(a, w) = \text{argmax}_{\left\{\delta \geq 0,
\frac{g(a - \delta \mid w)}{g(a \mid w)} \leq M \right\}} \frac{g(a - \delta
\mid w)}{g(a \mid w)}$$ and
$$\delta_{\text{min}}(a, w) = \text{argmin}_{\left\{\delta \leq 0,
\frac{g(a - \delta \mid w)}{g(a \mid w)} \leq M \right\}} \frac{g(a - \delta
\mid w)}{g(a \mid w)}.$$

The above provides a strategy for implementing a shift at the level of a given
observation $(a_i, w_i)$, thereby allowing for all observations to be shifted to
an appropriate value -- whether $\delta_{\text{min}}$, $\delta$, or
$\delta_{\text{max}}$.

### _Interlude:_ Constructing Optimal Stacked Regressions with `sl3`

To easily incorporate ensemble machine learning into the estimation procedure,
we rely on the facilities provided in the [`sl3` R
package](https://sl3.tlverse.org). For a complete guide on using the `sl3` R
package, consider consulting https://sl3.tlverse.org, or https://tlverse.org for
the [`tlverse` ecosystem](https://github.com/tlverse), of which `sl3` is a major
part.

Using the framework provided by the [`sl3` package](https://sl3.tlverse.org),
the nuisance parameters of the TML estimator may be fit with ensemble learning,
using the cross-validation framework of the Super Learner algorithm of
@vdl2007super.

```{r sl3_lrnrs}
# learners used for outcome regression
lrn1 <- Lrnr_mean$new()
lrn2 <- Lrnr_glm$new()
lrn3 <- Lrnr_ranger$new()
sl_lrn <- Lrnr_sl$new(
  learners = list(lrn1, lrn2, lrn3),
  metalearner = Lrnr_nnls$new()
)

# learners used for intervention conditional density estimatoin
lrn1_dens <- Lrnr_condensier$new(
  nbins = 25, bin_estimator = lrn2,
  bin_method = "dhist"
)
lrn2_dens <- Lrnr_condensier$new(
  nbins = 15, bin_estimator = lrn3,
  bin_method = "dhist"
)
lrn3_dens <- Lrnr_condensier$new(
  nbins = 20, bin_estimator = lrn2,
  bin_method = "dhist"
)
lrn4_dens <- Lrnr_condensier$new(
  nbins = 40, bin_estimator = lrn1,
  bin_method = "dhist"
)
sl_lrn_dens <- Lrnr_sl$new(
  learners = list(lrn1_dens, lrn2_dens, lrn3_dens, lrn4_dens),
  metalearner = Lrnr_solnp_density$new()
)
```

As seen above, we can generate two different ensemble learners for the two
nuisance regressions that must be fit in the process of computing this TML
estimator. In particular, we use a Super Learner composed of an intercept model,
a GLM, and a random forest (as implemented in [the `ranger` R
package](cran.r-project.org/package=ranger)) for fitting the outcome regressions
(often denoted "Q" in the literature) while we use variations of these learners,
through [the `condensier` R package](https://github.com/osofr/condensier), for
the conditional density estimation needed in fitting the treatment mechanism
(often denoted "g" in the literature).

We make the above explicit with respect to standard notation by bundling the
ensemble learners into a list object below:

```{r make_lrnr_list}
# specify outcome and treatment regressions and create learner list
Q_learner <- sl_lrn
g_learner <- sl_lrn_dens
learner_list <- list(Y = Q_learner, A = g_learner)
```

The `learner_list` object above specifies the role that each of the ensemble
learners we've generated is to play in computing initial estimators to be used
in building a TMLE for the parameter of interest here. In particular, it makes
explicit the fact that our `Q_learner` is used in fitting the outcome regression
while our `g_learner` is used in fitting our treatment mechanism regression.

### Initializing `vimshift` through its `tmle3_Spec`

To start, we will initialize a specification for the TMLE of our parameter of
interest (called a `tmle3_Spec` in the `tlverse` nomenclature) simply by calling
`tmle_shift`. We specify the argument `shift_grid = seq(-1, 1, by = 0.5)`
when initializing the `tmle3_Spec` object to communicate that we're interested
in assessing the mean counterfactual outcome over a grid of shifts `r seq(-1,
1, by = 0.5)` on the scale of the treatment $A$ (note that the numerical
choice of shift is an arbitrarily chosen set of values for this example).

```{r vim_spec_init}
# what's the grid of shifts we wish to consider?
delta_grid <- seq(-1, 1, 0.5)

# initialize a tmle specification
tmle_spec <- tmle_vimshift_delta(shift_grid = delta_grid,
                                 max_shifted_ratio = 2)
```

As seen above, the `tmle_vimshift` specification object (like all `tmle3_Spec`
objects) does _not_ store the data for our specific analysis of interest. Later,
we'll see that passing a data object directly to the `tmle3` wrapper function,
alongside the instantiated `tmle_spec`, will serve to construct a `tmle3_Task`
object internally (see the `tmle3` documentation for details).


### Targeted Estimation of Stochastic Interventions Effects

One may walk through the step-by-step procedure for  fitting the TML estimator
of the mean counterfactual outcome under each shift in the grid, using the
machinery exposed by the [`tmle3` R package](https://tmle3.tlverse.org/) (see
below); however, the step-by-step procedure is often not of interest.

```{r fit_tmle_manual, eval=FALSE}
# NOT RUN -- SEE NEXT CODE CHUNK

# define data (from tmle3_Spec base class)
tmle_task <- tmle_spec$make_tmle_task(data, node_list)

# define likelihood (from tmle3_Spec base class)
likelihood_init <- tmle_spec$make_initial_likelihood(tmle_task, learner_list)

# define update method (fluctuation submodel and loss function)
updater <- tmle_spec$make_updater()
likelihood_targeted <- Targeted_Likelihood$new(likelihood_init, updater)

# invoke params specified in spec
tmle_params <- tmle_spec$make_params(tmle_task, likelihood_targeted)
updater$tmle_params <- tmle_params

# fit TML estimator update
tmle_fit <- fit_tmle3(tmle_task, likelihood_targeted, tmle_params, updater)

# extract results from tmle3_Fit object
tmle_fit
```

Instead, one may invoke the `tmle3` wrapper function (a user-facing convenience
utility) to fit the series of TML estimators (one for each parameter defined by
the grid delta) in a single function call:

```{r fit_tmle_auto}
# fit the TML estimator
tmle_fit <- tmle3(tmle_spec, data, node_list, learner_list)
tmle_fit
```

_Remark_: The `print` method of the resultant `tmle_fit` object conveniently
displays the results from computing our TML estimator.


### Inference with Marginal Structural Models

In the directly preceding section, we consider estimating the mean
counterfactual outcome $\psi_n$ under several values of the intervention
$\delta$, taken from the aforementioned $\delta$-grid. We now turn our attention
to an approach for obtaining inference on a single summary measure of these
estimated quantities. In particular, we propose summarizing the estimates
$\psi_n$ through a marginal structural model (MSM), obtaining inference by way
of a hypothesis test on a parameter of this working MSM. For a data structure
$O = (W, A, Y)$, let $\psi_{\delta}(P_0)$ be the mean outcome under a shift
$\delta$ of the treatment, so that we have $\vec{\psi}_{\delta} =
(\psi_{\delta}: \delta)$ with corresponding estimators $\vec{\psi}_{n, \delta}
= (\psi_{n, \delta}: \delta)$. Further, let $\beta(\vec{\psi}_{\delta}) =
\phi((\psi_{\delta}: \delta))$.

For a given MSM $m_{\beta}(\delta)$, we have that
$$\beta_0 = \text{argmin}_{\beta} \sum_{\delta}(\psi_{\delta}(P_0) -
m_{\beta}(\delta))^2 h(\delta),$$
which is the solution to
$$u(\beta, (\psi_{\delta}: \delta)) = \sum_{\delta}h(\delta)
\left(\psi_{\delta}(P_0) - m_{\beta}(\delta) \right) \frac{d}{d\beta}
m_{\beta}(\delta) = 0.$$
This then leads to the following expansion
$$\beta(\vec{\psi}_n) - \beta(\vec{\psi}_0) \approx -\frac{d}{d\beta} u(\beta_0,
\vec{\psi}_0)^{-1} \frac{d}{d\psi} u(\beta_0, \psi_0)(\vec{\psi}_n -
\vec{\psi}_0),$$
where we have
$$\frac{d}{d\beta} u(\beta, \psi) = -\sum_{\delta} h(\delta) \frac{d}{d\beta}
m_{\beta}(\delta)^t \frac{d}{d\beta} m_{\beta}(\delta)
-\sum_{\delta} h(\delta) m_{\beta}(\delta) \frac{d^2}{d\beta^2}
m_{\beta}(\delta),$$
which, in the case of an MSM that is a linear model (since
$\frac{d^2}{d\beta^2} m_{\beta}(\delta) = 0$), reduces simply to
$$\frac{d}{d\beta} u(\beta, \psi) = -\sum_{\delta} h(\delta) \frac{d}{d\beta}
m_{\beta}(\delta)^t \frac{d}{d\beta} m_{\beta}(\delta),$$
and
$$\frac{d}{d\psi}u(\beta, \psi)(\psi_n - \psi_0) = \sum_{\delta} h(\delta)
\frac{d}{d\beta} m_{\beta}(\delta) (\psi_n - \psi_0)(\delta),$$
which we may write in terms of the efficient influence function (EIF) of $\psi$
by using the first order approximation $(\psi_n - \psi_0)(\delta) =
\frac{1}{n}\sum_{i = 1}^n \text{EIF}_{\psi_{\delta}}(O_i)$,
where $\text{EIF}_{\psi_{\delta}}$ is the efficient influence function (EIF) of
$\vec{\psi}$.

Now, say, $\vec{\psi} = (\psi(\delta): \delta)$ is d-dimensional, then we may
write the efficient influence function of the MSM parameter $\beta$ (assuming a
linear MSM) as follows
$$\text{EIF}_{\beta}(O) = \left(\sum_{\delta} h(\delta) \frac{d}{d\beta}
m_{\beta}(\delta) \frac{d}{d\beta} m_{\beta}(\delta)^t \right)^{-1} \cdot
\sum_{\delta} h(\delta) \frac{d}{d\beta} m_{\beta}(\delta)
\text{EIF}_{\psi_{\delta}}(O),$$ where the first term is of dimension
$d \times d$ and the second term is of dimension $d \times 1$.

In an effort to generalize still further, consider the case where
$\psi_{\delta}(P_0) \in (0, 1)$ -- that is, $\psi_{\delta}(P_0)$ corresponds
to the probability of some event of interest. In such a case, it would be more
natural to consider a logistic MSM
$$m_{\beta}(\delta) = \frac{1}{1 + \exp(-f_{\beta}(\delta))},$$
where $f_{\beta}$ is taken to be linear in $\beta$ (e.g.,
$f_{\beta} = \beta_0 + \beta_1 \delta + \ldots$). In such a case, we have the
parameter of interest
$$\beta_0 = \text{argmax}_{\beta} \sum_{\delta} \left(\psi_{\delta}(P_0)
\text{log} m_{\beta}(\delta) + (1 - \psi_{\delta}(P_0))\log(1 -
m_{\beta}(\delta))\right)h(\delta),$$
where $\beta_0$ solves the following
$$
\sum_{\delta} h(\delta) \frac{d}{d\beta} f_{\beta}(\delta) (\psi_{\delta}(P_0)
- m_{\beta}(\delta)) = 0.$$

Inference from a working MSM is rather straightforward. To wit, the limiting
distribution for $m_{\beta}(\delta)$ may be expressed
$$\sqrt{n}(\beta_n - \beta_0) \to N(0, \Sigma),$$
where $\Sigma$ is the empirical covariance matrix of $\text{EIF}_{\beta}(O)$.

```{r msm_fit}
tmle_fit$summary[6:7,]
```

#### Directly Targeting the MSM Parameter $\beta$

Note that in the above, a working MSM is fit to the individual TML estimates of
the mean counterfactual outcome under a given value of the shift $\delta$ in the
supplied grid. The parameter of interest $\beta$ of the MSM is asymptotically
linear (and, in fact, a TML estimator) as a consequence of its construction from
individual TML estimators. In smaller samples, it may be prudent to perform a
TML estimation procedure that targets the parameter $\beta$ directly, as opposed
to constructing it from several independently targeted TML estimates. An
approach for constructing such an estimator is proposed in the sequel.

Let $C = \left(\sum_{\delta} h(\delta) \frac{d}{d\beta} m_{\beta}(\delta)
\frac{d}{d\beta} m_{\beta}(\delta)^t \right)$, then
$$\text{EIF}_{\beta}(O) = C^{-1} \cdot \sum_{\delta} h(\delta)
\frac{d}{d\beta} m_{\beta}(\delta)(Y - \overline{Q}(A,W) + C^{-1} \sum_{\delta}
h(\delta) \frac{d}{d\beta} m_{\beta}(\delta) \left(\int \overline{Q}(a,w)
g_{\delta}^0(a \mid w) - \Psi_{\delta}\right).$$

Suppose a simple working MSM $\mathbb{E}Y_{g^0_{\delta}} = \beta_0 + \beta_1
\delta$, then a TML estimator targeting $\beta_0$ and $\beta_1$ may be
constructed as
$$\overline{Q}_{n, \epsilon}(A,W) = \overline{Q}_n(A,W) + \epsilon (H_1(g),
H_2(g),$$ for all $\delta$, where $H_1(g)$ is the auxiliary covariate for
$\beta_0$ and $H_2(g)$ is the auxiliary covariate for $\beta_1$.

To construct a targeted maximum likelihood estimator that directly targets the
parameters of the working marginal structural model, we may use the
`tmle_vimshift_msm` Spec (instead of the `tmle_vimshift_delta` Spec that
appears above):

```{r vim_targeted_msm_fit}
# what's the grid of shifts we wish to consider?
delta_grid <- seq(-1, 1, 0.5)

# initialize a tmle specification
tmle_msm_spec <- tmle_vimshift_msm(shift_grid = delta_grid,
                                   max_shifted_ratio = 2)

# fit the TML estimator and examine the results
tmle_msm_fit <- tmle3(tmle_msm_spec, data, node_list, learner_list)
tmle_msm_fit
```

## References

